---
title: "DS3 projects"
author: "Matthew Munoz and Fabio Pecora"
output: html_document
date: "2024-06-24"
---

```{r}
library(tidyverse)
library(zoo)
library(stm)
```


```{r download csv files}
report <- read.csv("google_trends_report.csv")
crime <- read.csv("crime.csv")
welfare <- read.csv("google_trends_welfare.csv")
```

```{r adding easier to understand colummns}
colnames(report) <- c("y", "m" ,"trends")
colnames(crime) <- c("y", "m","trends")
colnames(welfare) <- c("y", "m","trends")
```

```{r only for the og dataset}
report$ym <- with(report, sprintf("%d-%02d", y, m))
crime$ym <- with(crime, sprintf("%d-%02d", y, m))
welfare$ym <- with(welfare, sprintf("%d-%02d", y, m))


report <- report |> select(ym, trends)
crime <- crime |> select(ym, trends)
welfare <- welfare |> select(ym, trends)
```



```{r function to change year-month to year-month-date}
add_date <- function(df) {
  df <- df |>  
    mutate(ymd=as.Date(paste(ym, "-20", sep=""))) |> 
    select(ymd, trends)
  }
```

```{r add date to data frames}
report <- add_date(report) 
crime <- add_date(crime) 
welfare <- add_date(welfare) 
```


```{r function to add president column}
add_pres <- function(df) {
  bush_end <- as.Date("2009-01-20")
  trump_start <- as.Date("2017-01-20")
  
  df$president <- 
    ifelse(df$ymd >= bush_end & df$ymd < trump_start, "obama", 
           ifelse(df$ymd < bush_end, "bush", "trump")) 
  
  return(df)
}
```

```{r add president column to each vector}
report <- add_pres(report)
crime <- add_pres(crime) 
welfare <- add_pres(welfare)
```

```{r function to add zeros and ones}
zero_or_one <- function(df) {
  df$bush <- 
    ifelse(df$president == "bush", 1, 0) 
  df$obama <- 
    ifelse(df$president == "obama", 1, 0) 
  df$trump <- 
    ifelse(df$president == "trump", 1, 0) 
  
  return(df)
}
```

```{r add president column to each vector}
report <- zero_or_one(report)
crime <- zero_or_one(crime) 
welfare <- zero_or_one(welfare)
```

```{r function to plot figure 4}
plot_4 <- function(df, start, end) {
  df |>
    ggplot(aes(x = ymd, y = trends, color = president)) +
    geom_point(shape = 1) +
    geom_smooth(method = "lm", se = FALSE) + 
    ylim(start, end) +
    xlab("Year") +
    ylab("Google Trends")
}
```

```{r trends plot}
plot_4(report, 0, 100)
plot_4(crime, 0, 100)
plot_4(welfare, 0, 100)
```



```{r regression model}
fit.report <- lm(trends~ymd+bush+trump+obama, data=report |> filter(trends > 0))
fit.crime <- lm(trends~ymd+bush+trump+obama, data=crime |> filter(trends > 0))
fit.welfare <- lm(trends~ymd+bush+trump+obama, data=welfare |> filter(trends > 0))
summary(fit.report)
summary(fit.crime)
summary(fit.welfare)
```


```{r topic model}
load("TopicModel.RData")
document_topics <- make.dt(immigrFit, meta = out$meta)
topic_terms <- t(exp(immigrFit$beta$logbeta[[1]]))
rownames(topic_terms) <- out$vocab
colnames(topic_terms) <- sprintf("Topic%d", 1:ncol(topic_terms))
```

```{r prep for figures 2 and 3}
document_topics <- document_topics |> 
   mutate(date = as.Date(date, format = "%Y-%m-%d"),yearmonth = ym(format(date, "%Y-%m")))

time_line <- document_topics |> 
  select(time, yearmonth) |> 
  group_by(time) |> 
  summarize(yearmonth = min(yearmonth))

date1 <- as.Date("2015-06-01")
date2 <- as.Date("2017-01-01")

my_colors <- c("magenta", "red3", "blue")
```

```{r creating figure 2}
plot_2_table <- document_topics |> 
  select(channel, duration, yearmonth, time) |> group_by(yearmonth, channel, time) |>
  summarize(summation=sum(duration)) |> arrange(summation) 
 

plot_2_table |> ungroup() |> 
  group_by(channel, time) |> 
  ggplot(aes(x = yearmonth, y = summation, color = channel, group = interaction(channel, time))) + 
  geom_vline(xintercept = as.numeric(date1), linetype = "dashed", color = "black") +
  geom_vline(xintercept = as.numeric(date2), linetype = "dashed", color = "black") +
  geom_point() + geom_smooth( se=FALSE) + scale_color_manual(values = my_colors) +
  theme(axis.title.x = element_blank(), legend.position = "bottom", legend.title = element_blank()) +
  ylab("Num Monthly\nImmigration Segs") 
``````

```{r creating figure 3a(crime)}
plot_3a_table <- document_topics |> 
  select(Topic1, Topic3, channel, duration, yearmonth, time) |> group_by(yearmonth, channel, time) |>
  summarize(summation=sum(duration*Topic1 + duration*Topic3)) |> arrange(summation) 
 

plot_3a_table |> ungroup() |> 
  group_by(channel, time) |> 
  ggplot(aes(x = yearmonth, y = summation, color = channel, group = interaction(channel, time))) + 
  geom_vline(xintercept = as.numeric(date1), linetype = "dashed", color = "black") +
   geom_vline(xintercept = as.numeric(date2), linetype = "dashed", color = "black") +
  geom_point() + geom_smooth( se=FALSE) + scale_color_manual(values = my_colors) +
  theme(axis.title.x = element_blank(), legend.position = "bottom", legend.title = element_blank()) +
  ylab("Immiger + Crime\nNews Coverage") 
``````


```{r creating figure 3b(welfare)}
plot_3b_table <- document_topics |> 
  select(Topic13, channel, duration, yearmonth, time) |> group_by(yearmonth, channel, time) |>
  summarize(summation=sum(duration*Topic13)) |> arrange(summation) 

plot_3b_table |> ungroup() |> 
  group_by(channel, time) |> 
  ggplot(aes(x = yearmonth, y = summation, color = channel, group = interaction(channel, time))) + 
  geom_vline(xintercept = as.numeric(date1), linetype = "dashed", color = "black") +
   geom_vline(xintercept = as.numeric(date2), linetype = "dashed", color = "black") +
  geom_point() + geom_smooth( se=FALSE) + scale_color_manual(values = my_colors) +
    theme(axis.title.x = element_blank(), legend.position = "bottom", legend.title = element_blank()) +
  ylab("Immiger Welfare\nNew Coverage") 
``````





